---
title: "CORE S115 Fall 2026: Lab 2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE)
```


## Dataframes

Dataframes in R are tables where each column is a collection of values. Each row can be thought of as a "data point". For example, the 1st value of all the collections become the 1st "data point". The 10th value of all the collections become the 10th data point, and so on. 

The following code creates three different collections and puts them together in a data frame. 

```{r}

## Data taken from https://www.billboard.com/charts/hot-100/ week of September 5

s = c("Choosin' Texas", 'Boston', 'I knew It, I knew You', 'Been By Now', 'Hate That I Made You Love Me', 'Dracula',
      'So Easy (To Fall in Love)', "I Can't Love You Anymore", 'Man I need', 'Be Her')

a = c('Ella Langley', 'Stella Lefty', 'Taylor Swift', 'Morgan Wallen', 'Ariana Grande', 'Tame Impala & JENNIE',
      'Olivia Dean', 'Ella Langley & Morgan Wallen', 'Olivia Dean', 'Ella Langley')

w = c(45,21,12,5,13,48,48,18,53,28)

billboard = data.frame(
  song = s,  # assigning collection s to column called song
  artist = a, # assigning collection a to column called artist
  weeks_on_chart = w # assigning collection w to a column called weeks_on_chart
)

```


#### Written Q1
How many rows and columns do you think the dataframe billboard has? 

**Answer**

The following functions can help verify your answers

```{r}

print(length(billboard))
print(nrow(billboard))
print(colnames(billboard))

```

#### Written Q2
What do you think the functions `length`,`nrow` and `colnames` do? 

**Answer**


## Accessing columns from dataframes

You can access columns from a dataframe by using the "$" symbol. For example, the code below prints the values in the column `song`. 

```{r}

print(billboard$song)

```


#### Coding Q1

Print the values in the column `artist` in the dataframe `billboard`

```{r}

## WRITE YOUR CODE BELOW THIS LINE

```


#### Coding Q2

Assign the maximum value in the column `weeks_on_chart` in `billboard` to a variable called `maximum`. Print "The maximum weeks on chart is [x]" where [x] is replaced by the value stored in the variable `maximum`. Make sure to use the variable in your print statement! 

*Hint: Look at the paste() function from last class*

```{r}

## WRITE YOUR CODE BELOW THIS LINE

```


#### Coding Q3

Assign the average value in the column `weeks_on_chart` in `billboard` to a variable called `average`. Print "The difference between the maximum and average weeks on chart is [x]" where [x] is replaced by the difference between the values stored in the variables `maximum` and `average`. Make sure to use these two variables in your print statement! 

```{r}
## WRITE YOUR CODE BELOW THIS LINE

```


## Adding columns to dataframes

You can also add columns to existing dataframes by using the "$" sign. For example, the following code adds a column called rank. 

```{r}
billboard$rank = c(1:length(billboard))
```


#### Written Q3
Explain why the following code causes an error. 

**Answer**

```{r, error=TRUE}

billboard$rank2 = c(1:13)

```


#### Coding Q4

Assign the average value in the column `rank` in `billboard` to a variable called `average`. Print "The difference between the maximum and average weeks on chart is [x]" where [x] is replaced by the difference between the values stored in the variables `maximum` and `average` for each row. Make sure to use these two variables in your print statement! 

```{r}
## WRITE YOUR CODE BELOW THIS LINE

```


You can also use values of existing columns to create new columns. For example, the following code adds two to each value of rank, and stores it in a variable called `rank2`

```{r}

billboard$rank2 = billboard$rank + 2

```

#### Coding Q5

Create a variable called `days_on_chart` that stores the number of days that each song has been on the billboard.

*Hint: think about how you can convert weeks into days*

```{r}
## WRITE YOUR CODE BELOW THIS LINE

```


## Using conditionals in dataframes

You can use conditionals to create new columns. When you do this, the conditional gets applied to every individual value in the column.

For example, the following code creates a column called `multiple_artists` which holds TRUE if the artist name has the string '&' in it, showing there is more than one artist.

(Note the string '&' is different from the logical operator `&` which we will use later!!)

```{r}

billboard$colab = ifelse('&' %in% billboard$artist, 'Is colab', 'Is not colab')

```


#### Coding Q6

Create a variable called `weeks_category` that stores the string '>=40' if a song has been on the billboard for 40 weeks or more, '>=10' if a song has been on the billboard for 10 weeks or more (but less than 20 weeks), and '< 10' otherwise. 

*Hint: Look up the syntax for nested conditionals from last class*

```{r}
## WRITE YOUR CODE BELOW THIS LINE

```


## Creating subsets of dataframes

Sometimes you want to look at only subsets of your dataframe. For example, you might want to look at only the songs which are NOT by Olivia Dean. You can use the `subset()` function for that. 

```{r}

not_olivia = subset(billboard, artist != 'Olivia Dean')

```

You can also combine multiple conditions using `and` (&) or `or` (|)

```{r}
x = subset(billboard, artist != 'Sabrina Carpenter' & weeks_on_chart >=15)
y = subset(billboard, artist != 'Sabrina Carpenter' | weeks_on_chart < 10)

```

#### Written Q4
Describe what is stored in x and y, and explain why that is the case. 

**Answer**

#### Coding Q7

Find the subset of the songs in the billboard dataframe which have either been on the billboard for greater than 15 weeks, or been on the billboard for less than 5 weeks. Assign this to a variable called `billboard_q7`

```{r}
## WRITE YOUR CODE BELOW THIS LINE

```



#### Coding Q8

Find the subset of the songs in the billboard dataframe which have been on the billboard for greater than 15 weeks and have been on the billboard for less than 5 weeks. Assign this subset to a variable called `billboard_q8`

```{r}
## WRITE YOUR CODE BELOW THIS LINE

```

#### Written Q5

How many rows did the dataframe from Q8 have? Why does this make sense? 

**Answer**

#### Coding Q9

Find the subset of the songs in the billboard dataframe which Ella Langley contributed to (either as a single artist or with someone else) and which has been on the billboard for less than 30 weeks. Then print the names of the songs in this subset. 

```{r}
## WRITE YOUR CODE BELOW THIS LINE

```


#### Coding Q10

Assign the average value in the column `weeks_on_chart` in the subset `billboard_q7` to a variable called `average`. Print "The difference between the maximum and average weeks on chart is [x]" where [x] is replaced by the difference between the values stored in the variables `maximum` and `average`. 
 

```{r}
## WRITE YOUR CODE BELOW THIS LINE

```


## Putting it together (Homework)

Using a combination of code and text, do the following: 

1. Load the titanic dataset. 

2. Describe the titanic dataset: how many rows, how many columns, what are the columns?

3. Investigate whether the number of women who survived is greater than the number of men who survived. 

4. Investigate whether the proportion of women who survived is greater than the proportion of men who survived. 

5. What are the average ages of the people who survived and did not survive? **Hint: you will notice that there are some NA values in the data which will make the entire average NA. To avoid this, use na.rm=TRUE). 

6. What are the average fares of the people who survived and did not survive? **Hint: you will notice that there are some NA values in the data which will make the entire average NA. To avoid this, use na.rm=TRUE). 

7. Based on your investigation above, discuss what factors you think might have affected a person's chance of survival, and what factors were not important. 
